---
title: "BDA problem set"
author: "Pradip Hazra (413)"
date: "2025-04-08"
output: word_document
---

##### 1. Suppose you have observed a coin flip experiment with the following results: 12 heads and 8 tails. Assume a Beta(2,2) prior for the probability of heads. Use the Beta-Binomial model to obtain the posterior distribution, posterior mean, and a 95% credible interval for the probability of heads.

```{r}
# Prior parameters for Beta distribution
alpha_prior <- 2
beta_prior <- 2

# Observed data (12 heads, 8 tails)
x <- 12  # number of heads
n <- 20  # total flips (12 heads + 8 tails)

# Posterior parameters for Beta distribution
alpha_post <- alpha_prior + x
beta_post <- beta_prior + (n - x)

# Display the posterior parameters
cat("Posterior parameters:\n")
cat("Alpha: ", alpha_post, "\n")
cat("Beta: ", beta_post, "\n")

# Posterior mean
posterior_mean <- alpha_post / (alpha_post + beta_post)
cat("Posterior mean: ", posterior_mean, "\n")

# 95% credible interval (2.5th and 97.5th percentiles)
credible_interval <- qbeta(c(0.025, 0.975), alpha_post, beta_post)
cat("95% credible interval: [", credible_interval[1], ", ", credible_interval[2], "]\n")

# Optional: plot the posterior Beta distribution using plot function
x_values <- seq(0, 1, length.out = 200) # Generate a sequence of x values
y_values <- dbeta(x_values, alpha_post, beta_post) # Calculate the density for each x

plot(x_values, y_values,
     type = "l", # "l" for line plot
     main = "Posterior Beta Distribution",
     xlab = "Probability of heads (θ)",
     ylab = "Density")

# Add posterior mean line
abline(v = posterior_mean, col = "red", lty = 2)

# Add credible interval lines
abline(v = credible_interval[1], col = "blue", lty = 3)
abline(v = credible_interval[2], col = "blue", lty = 3)

# Add legend
legend("topright",
       legend = c("Posterior Density", "Posterior Mean", "95% Credible Interval"),
       col = c("black", "red", "blue"),
       lty = c(1, 2, 3))
```
##### 4. A call center records customer complaints daily:
(3,2,5,1,4,3,2,6,5,2)(3, 2, 5, 1, 4, 3, 2, 6, 5, 2)
Assume a Gamma(2,1) prior for the complaint rate. Compute the posterior distribution, posterior mean, and a 95% credible interval for the complaint rate.

```{r}
# Given data
complaints <- c(3, 2, 5, 1, 4, 3, 2, 6, 5, 2)

# Prior parameters (Gamma(2, 1))
alpha_0 <- 2
beta_0 <- 1

# Sufficient statistics (sum of complaints and number of observations)
n <- length(complaints)
sum_complaints <- sum(complaints)

# Posterior parameters
alpha_post <- alpha_0 + sum_complaints
beta_post <- beta_0 + n

# Posterior mean
posterior_mean <- alpha_post / beta_post

# 95% credible interval (using quantile function of the Gamma distribution)
credible_interval <- qgamma(c(0.025, 0.975), shape = alpha_post, rate = beta_post)

# Results
cat("Posterior mean: ", posterior_mean, "\n")
cat("95% credible interval: (", credible_interval[1], ", ", credible_interval[2], ")\n")

# Plot the posterior distribution
# Define a range of lambda values for plotting
lambda_values <- seq(0, 10, by = 0.01)

# Compute the Gamma density for each lambda value
posterior_density <- dgamma(lambda_values, shape = alpha_post, rate = beta_post)

# Plot the posterior distribution
plot(lambda_values, posterior_density, type = "l", lwd = 2, col = "blue", 
     main = "Posterior Distribution of Complaint Rate (Lambda)",
     xlab = "Complaint Rate (Lambda)", ylab = "Density", 
     ylim = c(0, max(posterior_density) * 1.1))
abline(v = posterior_mean, col = "red", lty = 2) # Add vertical line for posterior mean
abline(v = credible_interval, col = "green", lty = 2) # Add vertical lines for credible interval

# Add a legend
legend("topright", legend = c("Posterior Density", 
                              paste("Posterior Mean =", round(posterior_mean, 2)), 
                              paste("95% CI: (", round(credible_interval[1], 2), ",", round(credible_interval[2], 2), ")")),
       col = c("blue", "red", "green"), lty = c(1, 2, 2), lwd = c(2, 1, 1))

```

##### 7. Given a dataset of student test scores:
(56,60,58,62,59,55,61,63,57,60) 
Assume a Normal prior N(60, ) and a known variance of 4. Compute the posterior distribution for the mean, the posterior mean, and a 95% credible interval.

```{r}
# Given data
data <- c(56, 60, 58, 62, 59, 55, 61, 63, 57, 60)
n <- length(data)
mean_data <- mean(data)

# Prior parameters
prior_mean <- 60
prior_variance <- 4
prior_sd <- sqrt(prior_variance)

# Known variance of the data distribution
likelihood_variance <- 4
likelihood_sd <- sqrt(likelihood_variance)

# Calculate posterior parameters for the mean
posterior_mean <- (prior_mean * (n / likelihood_variance) + mean_data * (1 / prior_variance)) /
                  ((n / likelihood_variance) + (1 / prior_variance))
posterior_variance <- 1 / ((n / likelihood_variance) + (1 / prior_variance))
posterior_sd <- sqrt(posterior_variance)

# Calculate the 95% credible interval
alpha <- 0.05
credible_interval <- posterior_mean + c(-1, 1) * qnorm(1 - alpha / 2) * posterior_sd

# R code to draw only the posterior curve

# Define a range of values for the mean around the posterior mean
mu_values_posterior <- seq(posterior_mean - 3 * posterior_sd, posterior_mean + 3 * posterior_sd, length.out = 200)

# Calculate posterior density
posterior_density <- dnorm(mu_values_posterior, mean = posterior_mean, sd = posterior_sd)

# Plot the posterior curve
plot(mu_values_posterior, posterior_density, type = "l", col = "red",
     xlab = "Mean (μ)", ylab = "Posterior Density",
     main = "Posterior Distribution for the Mean")

# Add a vertical line for the posterior mean
abline(v = posterior_mean, col = "red", lty = 2)

# Add vertical lines for the credible interval
abline(v = credible_interval[1], col = "red", lty = 3)
abline(v = credible_interval[2], col = "red", lty = 3)

# Add a legend
legend("topright",
       legend = c("Posterior Density", "Posterior Mean", "95% Credible Interval"),
       col = c("red", "red", "red"),
       lty = c(1, 2, 3))
```



